Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value

### 2021-09-15 docs/accepted/manuscript-accepted.Rmd
add: new directory for accepted version in Meta-Psychology

affiliation:
  - id            : "1"
    institution   : "School of Human Kinetics University of Ottawa"
  - id            : "2"
    institution   : "Department of Kinesiology McMaster University"

authornote: |
  Submitted to *Meta-Psychology*. Participate in open peer review by commenting through hypothes.is directly on this preprint. The full editorial process of articles under review at *Meta-Psychology* can be found here: https://tinyurl.com/mp-submissions. You will find this preprint by searching for the first authors name.

@@

abstract: |
  The self-controlled motor learning literature consists of experiments that compare a group of learners who are provided with a choice over an aspect of their practice environment to a group who are yoked to those choices. A qualitative review of the literature suggests an unambiguous benefit from self-controlled practice. A meta-analysis was conducted on the effects of self-controlled practice on retention test performance measures with a focus on assessing and potentially correcting for selection bias in the literature, such as publication bias and *p*-hacking. First, a naïve random effects model was fit to the data and a moderate benefit of self-controlled practice, $g = .44\,(k = 52, N = 3134, 95\%\,CI\,[.31,\,.56])$, was found. Second, publication status was added to the model as a potential moderator, revealing a significant difference between published and unpublished findings, with only the former reporting a benefit of self-controlled practice. Third, to investigate and adjust for the impact of selectively reporting statistically significant results, a weight-function model was fit to the data with a one-tailed *p*-value cutpoint of .025. The weight-function model revealed substantial selection bias and estimated the true average effect of self-controlled practice as $g = .107\,(95\%\,CI\,[.047,\,.18])$. *P*-curve analyses were conducted on the statistically significant results published in the literature and the outcome suggested a lack of evidential value. Fourth, a suite of sensitivity analyses were conducted to evaluate the robustness of these results, all of which converged on trivially small effect estimates. Overall, our results suggest the benefit of self-controlled practice on motor learning is small and not currently distinguishable from zero.

@@

# NEED TO BE RUNNING THE DEVELOPMENT BRANCH VERSION OF PAPAJA
documentclass     : "apa7"
classoption       : "man" # manuscript output
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage{ragged2e}
  - \usepackage{pdflscape}
  - \pagewiselinenumbers
  - \raggedbottom
  #- \leftheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  #- \rightheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  #- \hypersetup{draft} # comment out after but right now will prevent error message about some link spanning multiple pages
  #- \usepackage{setspace}
  #- \AtBeginEnvironment{longtable}{\singlespacing}
  #- \AtBeginEnvironment{tablenotes}{\doublespacing}
  #- \captionsetup[table]{font={stretch=1.5}}
appendix:
  - "appendix_a.Rmd"
  - "appendix_b.Rmd"
---

@@

Figure \@ref(fig:fig1) shows that the number of experiments comparing self-controlled groups to yoked groups has been increasing since the original experiments by Janelle and his colleagues [-@Janelle1995-rj; -@Janelle1997-ht]. Researchers have experimented with giving learners control over a variety of variables in the practice environment. A qualitative assessment of the literature suggests that self-control is generally beneficial regardless of choice-type [@Wulf2016-gf]. For example, self-control has been effective when participants have been provided choice over what can be considered instructionally-relevant variables, such as: knowledge of results [@Patterson2010-tk], knowledge of performance [@Lim2015-qs], concurrent feedback [@Huet2009-wu], use of an assistive device [@Wulf2001-nb], observation of a skilled model [@Lemos2017-qx], practice schedule [@Wu2011-ha], practice volume [@Lessa2015-eq], and task difficulty [@Leiker2016-zd]. Additionally, self-controlled benefits have also been found for instructionally-irrelevant variables, such as: the colour of various objects in the practice environment [@Wulf2018-ie], other decorative choices [@Iwatsuki2019-tf], and the choice of what to do after the retention test is complete [@Lewthwaite2015-bd].

@@

In sum, the objectives of this meta-analysis were to estimate the true average effect of self-controlled learning and evaluate the evidential value of the self-controlled learning literature. Bias resulting from selective publication was addressed with weight function and *p*-curve models and effect size estimates were adjusted accordingly. A key theoretical question \textcolor{red}{related to the underlying mechanisms of putative self-controlled learning advantages (motivational versus informational influences)} was also addressed through moderator analyses, but, to anticipate, inferences will depend on the reliability of the evidence overall. Finally, sensitivity analyses were conducted in addition to pre-registered analyses in an effort to understand the extent that our conclusions depended on the modeling techniques and assumptions adopted.

@@

## Pre-registration
The procedures followed to conduct this meta-analysis were pre-registered and can be viewed at https://osf.io/qbg69 \hyperref[sec:sharing]{(see Data, materials, and code availability section)}. This meta-analysis was retrospective and earlier samples of the literature had been meta-analyzed prior to this pre-registration, albeit with different data collection procedures, scope, and excluding recent experiments. \textcolor{red}{This study adheres to PRISMA reporting guidelines} [@page2021].

## Literature search
The literature search and data extraction were conducted by three authors (BM, ZY, JH) and one research assistant (HS) independently. The goal of the search was to identify all articles that met the inclusion criteria for the meta-analysis. Specifically, randomized experiments were subject to five criteria for inclusion: 1) A self-control group in which participants were asked to make at least one choice during practice, 2) a yoked group that experienced the same practice conditions as the self-controlled group, 3) a delayed ~24-hour retention test or test with longer delay interval, 4) an objective measurement of motor performance, and 5) publication in a peer-reviewed journal or acceptance as part of a Master’s or PhD thesis. \textcolor{red}{The literature search was completed on August 2, 2019.}

The search commenced at PubMed and Google Scholar with the following query: *self-control** OR *self-regulat** OR *self-direct** OR *learner-control** OR *learner-regulat** OR *learner-direct** OR *subject-control** OR *subject-regulat** OR *subject-direct** OR *performer-control** OR *performer-regulat** OR *performer-direct** AND *motor learning**. The query retrieved 9014 hits on PubMed and 98,600 hits on Google Scholar. Each researcher excluded hits based on title alone or title and abstract when necessary, and quit searching the databases at self-selected intervals following extended periods of excluding 100\% of search results. Following an initial run of searching databases, each researcher employed their own search strategies, including reviewing the reference sections of reviews and included articles, consulting the "OPTIMAL" theory website[^1], and searching the ProQuest Thesis database.

@@

Many of the self-controlled learning experiments analyzed in this study included multiple dependent measures. However, including multiple measures from the same experiment introduces bias and inflates \textcolor{red}{Type 1 error} [@Scammacca2014-do]. Although there are a variety of methods for dealing with multiple measures from the same studies in meta-analysis, we chose to create a priority list and always selected the highest priority dependent measure that was reported. If the highest priority measure was not described in adequate detail to calculate the effect size, the authors were contacted and the data were requested. If the authors could not provide the data for the highest priority dependent measure reported in their study, the experiment was left out of our analysis.

@@

@@

## Data extraction
The four researchers separated into pairs and half of the included experiments \textcolor{red}{were coded independently by one pair. The other half were coded independently by the other pair.} The coding included varied moderators, publication year, and sample size. Also Hedges’ *g* was calculated from reported statistics and sample size using the `compute.es` package [@R-computees] in `R` [@R-base]. Effect sizes were calculated from means and standard deviations, test statistics like *t* and *F*, or from precisely reported *p*-values. When covariates were included in the analysis, the correlation coefficient for the covariate - dependent measure relationship was required to calculate accurate effect sizes. Since this information is often not reported, authors were contacted and the information was requested. One effect size was calculated for each of three time points for each experiment: Acquisition, retention, and transfer.

The independent data extractions were compared and inconsistent results were highlighted. There was 89\% absolute agreement between pairs of coders on 1344 data points. For those with disagreement, one of the researchers from the other coding pair reviewed the relevant experiment to confirm the value to be used in the analysis.[^3]

@@

In addition to extracting effect sizes, inferential statistics were scraped from published experiments that reported a statistically significant effect at retention. Two authors (BM and JH) independently completed a *p*-curve disclosure form consisting of a direct quote of the stated hypotheses for each experiment, the experimental design, and a direct quote of the results indicating a significant result **(see Appendix A)**. There was 94% absolute agreement between the independent forms. Mismatches were resolved with consensus.

@@

## \textcolor{red}{Risk of bias}
\textcolor{red}{All articles were assessed for risk of bias by the lead author using the Cochrane Risk of Bias 1.0 tool} [@higgins2011]. \textcolor{red}{Each article was coded as either high risk, unclear (some concerns), or low risk on 7 dimensions: sequence generation, allocation concealment, incomplete outcome data, selective outcome reporting, blinding of outcome assessment, blinding of participants and personnel, and other sources of bias.}

@@

## Adjusting for selection effects
Selection bias in the motor learning literature is likely caused by filtering based on the statistical significance of results [@Lohse2016-cf]. To assess and adjust for selection effects, the `R` package `weightr` [@Coburn2017-hv]  was used to fit a Vevea-Hedges weight function model to the retention data [@Vevea1995-uv]. The weight-function model estimates the true average effect, heterogeneity, and the probability that a non-significant result \textcolor{red}{survives censorship and is available for analysis}. Selection effects are modelled by a step function that divides the effects into two bins at one-tailed *p* =.025, coinciding with a two-tailed *p*-value of .05. The probability of a non-significant effect \textcolor{red}{surviving censorship to appear} in the model is estimated relative to the probability of observing a study with a significant effect. The selection-adjusted model was compared to the naïve random effects model with a likelihood ratio test. Better fit from an adjusted model suggests selection bias in the literature.

@@

Bias correction methods vary in their performance depending on the total amount of heterogeneity, the true average effect size, the amount of publication bias, and the intensity of *p*-hacking in the data [@Carter2019-vv]. To determine which bias correction models perform well in the various plausible conditions for data in this meta-analysis, model performance checks were conducted using the \href{http://www.shinyapps.org/apps/metaExplorer/}{Meta-Showdown Explorer} shiny app developed by Carter and colleagues [-@Carter2019-vv]. Simulated conditions were as follows: Medium publications bias (significant results published at 100\% frequency, non-significant published at 20\% frequency, wrong direction effects published at 5\% frequency), \textcolor{red}{medium questionable research practice environment (QRP; see Carter et al.} [-@Carter2019-vv] \textcolor{red}{for detailed explanation of QRP environment)}, $\tau$ = 0, .2; $g$ = 0, .2, .5; $k$ = 60, good performance defined as a maximum of .1 upward or downward bias, and maximum mean absolute error of .1, also tested with maximum bias and error values of .15. With good performance defined by a maximum bias in either direction of .1 and maximum absolute error of .1, the weight function model and, \textcolor{red}{to a lesser extent,} *p*-curve models provided coverage across all plausible conditions except the highest heterogeneity condition ($\tau$ = .4). With good performance defined as a maximum bias and error of .15, the precision-effect with standard error (PEESE) method provided good performance in all conditions. Therefore, sensitivity analyses were conducted on effect size data via *p*-curve and PEESE methods. \textcolor{red}{An additional sensitivity analysis of the estimated power among included studies was conducted with the \emph{z}-curve [@bartovs2020]. \emph{Z}-curve, like \emph{p}-curve, analyzes only statistically significant results and estimates the power of the included studies (called expected replication rate, ERR). However, unlike \emph{p}-curve, \emph{z}-curve is robust to heterogeneity because it fits a finite mixture model of seven distributions, allowing the underlying true effects to vary. Further, \emph{z}-curve also estimates the power of all studies that have been conducted (called expected discovery rate, EDR) which can be compared to the observed discovery rate in order to test for the presence of publication bias.}

@@

## \textcolor{red}{Risk of bias}
\textcolor{red}{The risk of bias assessment revealed lackluster reporting standards were pervasive among the included articles} (see Figure \@ref(fig:fig2)). \textcolor{red}{For example, comparing a self-control group to a yoked group usually involves first collecting a self-control participant, then their yoked counterpart. Despite this, most articles simply reported that the participants were randomly assigned to these conditions, with no indication of how this temporal constraint was addressed. A similar issue was observed with respect to addressing outliers and attrition. Over 75\% of the included articles failed to mention outliers and how they were addressed (captured by the incomplete outcome data dimension). Most studies included in this study were not double-blind, largely due to the inherent difficulties in conducting a double-blind study of self-controlled motor learning. While the risk of bias associated with a lack of double blinding has been debated} [see @howick2008], \textcolor{red}{it is nonetheless notable that double-blinding was rare among the included studies.}

(ref:fig2-caption) \textcolor{red}{Proportion of studies with low risk, some concerns, and high risk of bias in each of the seven dimensions of the Cochrane RoB 1.0 tool.}

@@

(ref:fig3-caption) Forest plot of Hedges' *g* (95% CI) for self-controlled versus yoked groups on retention tests. Size of squares is proportional to $1/\sigma^2$ (precision). Light grey polygons represent \textcolor{red}{95\% $CI$ estimates from publication-status moderator analysis. Estimates from unpublished studies center on $g$ = .003 and published studies on $g$ = .54}

```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center", out.height = "81%"}
knitr::include_graphics("../../figs/fig3.pdf")
```

@@

## Selection model
The weight-function model combines an effect size model and a selection model [@Hedges1996-yh]. The effect size model is equivalent to the naïve random effects model, specifying what the distribution of effect sizes would be in the absence of publication bias or other selection effects. The selection model accounts for the probability a given study survives selection based on its *p*-value and specifies how the effect size distribution is modified by selection. A weight-function model with a *p*-value cutpoint of (one-tailed) .025 was fit to the retention effect size estimates(see Figure \@ref(fig:fig4)). The results of a likelihood ratio test suggest the adjusted model was a significantly better fit to the data than the unadjusted model, $\chi^2$(*df* = 1) = 21.18, *p* < .0001.[^4] The adjusted effect size estimate was significantly different from zero, *g* = .107, *p* < .001, 95% *CI* [.05, .17]. According to the adjusted model, non-significant results were 6% as likely to survive selection as significant results. \textcolor{red}{Note that the \texttt{weightr} function failed to estimate the random effects model and the results reported here are based on a fixed-effect estimate.}

@@

(ref:fig4-caption) Funnel plot of self-controlled learning studies at retention. Standard error is plotted on the y-axis and Hedges' g is plotted on the x-axis. \textcolor{red}{Dark gray contour regions represent two-tailed p-values between .10 and .05 (not quite significant). The light gray contour regions represent two-tailed p-values between .05 and .01.} In the absence of bias (and other forms of heterogeneity), the most precise experiments would centre on the naïve random effects estimate near the \textcolor{red} {top} of the plot and as experiments get progressively less precise they would move \textcolor{red}{down} the plot and spread out symmetrically. In the presence of bias, one would expect experiments to \textcolor{red} cluster in the light gray contour regions. The clustering of experiments \textcolor{red}{in the positive light gray  contour region} in the above plot suggests substantial bias.

\vfill

@@

## *P*-curve
The purpose of the *p*-curve analysis was to investigate the evidential value in the published reports (*N* = 26) of statistically significant self-controlled learning benefits. Visual inspection of Figure \@ref(fig:fig4) reveals a v-shaped distribution with the greatest frequency of *p*-values in the <.05 bin. The observed *p*-curve was significantly flatter than would be expected if the experiments had 33% power, *p* = .0035, indicating an absence of evidential value. Conversely, the half *p*-curve [@Simonsohn2015-mn] was significantly right skewed, suggesting the presence of evidential value. Sensitivity analysis, however, revealed that the half curve does not remain significantly right skewed following removal of the most extreme *p*-value from the sample. \textcolor{red}{The estimated power of the included studies was 5\%, 95\% $CI$ [5\%, 17\%].}

## Interim discussion
The primary results described above suggest that selection effects have caused a seriously distorted record of self-controlled learning. Estimated benefits are less than one third of the naïve estimate, *g* = .107, 95% *CI* \textcolor{red}{[.05, .17]}. The *p*-curve analysis failed to detect robust evidence of a self-controlled learning effect. The performance of the weight-function model depends on the specific conditions present in the meta-analysis, although these conditions are unknowable [@Carter2019-vv]. It was necessary to conduct sensitivity analyses with additional bias correction methods to assess the reliability of the selection-adjusted weight-function model estimate. Based on performance checks conducted under a range of plausible conditions, it was determined that sensitivity analyses conducted with a PEESE meta-regression and *p*-curve effect size estimation would provide good performance coverage across most plausible conditions.

@@

## Precision-effect with standard error (PEESE) model
When publication bias is present in a body of evidence, sample size and effect size can be negatively correlated [@Stanley2014-xl]. The PEESE model fits a quadratic relationship between effect size and standard error to reflect the intuition that publication bias is stronger for low precision studies than high precision studies. The rationale is that low precision studies need to overestimate effects to achieve significance and get published, while high precision studies can publish without exaggerated effects; thus, creating greater publication bias among lower precision studies [@Carter2019-vv; @Stanley2014-xl]. A weighted-least-squares regression model was fit with effect size regressed on the square of the standard error, weighted by the inverse of the variance:
$$g_{i} = b_{0} + b_{1}se_{i}^2 + e_{i}$$
The PEESE method estimated a non-significant benefit of self-controlled learning after controlling for publication bias, *g* = .054, *p* = .659.

## *P*-curve effect estimation
A *p*-curve model was fit to the overall retention effect size data, unlike the first primary *p*-curve which was fit to the reported significant results. The *p*-curve is a function of sample size and effect size, and because sample size is known, the effect size that provides the best fit to the observed *p*-curve can be estimated [@Simonsohn2014-xq]. A *p*-curve analysis conducted with th`R` package `dmetar` [@R-dmetar] was used to estimate the average effect size among the statistically significant effects in the meta-analysis. The model estimated an average effect of $g$ = .035.[^5] \textcolor{red}{The estimated power of included studies was 7\%, 95\% $CI$ [5\%, 22\%]. Unfortunately, \emph{p}-curve does not perform well in the presence of heterogeneity and these results should be interpreted cautiously.}

@@

## \textcolor{red}{\emph{Z}-curve}
\textcolor{red}{A \emph{z}-curve was fit to the overall retention data and estimated the power of statistically significant studies (ERR) as 14\%, 95\% CI [5\%, 35\%]. The power of all studies conducted (EDR) was estimated as 6\%, 95\% $CI$ [5\%, 14\%]. The 95\% confidence intervals for both the ERR and EDR failed to include the observed discovery rate of 48\%, suggesting significant publication bias in the data.}

>>>>

affiliation:
  - id            : "1"
    institution   : "School of Human Kinetics, University of Ottawa"
  - id            : "2"
    institution   : "Department of Kinesiology, McMaster University"

authornote: |
  \begin{center}
  In Press in Meta-Psychology
  \end{center}

@@

abstract: |
  The self-controlled motor learning literature consists of experiments that compare a group of learners who are provided with a choice over an aspect of their practice environment to a group who are yoked to those choices. A qualitative review of the literature suggests an unambiguous benefit from self-controlled practice. A meta-analysis was conducted on the effects of self-controlled practice on retention test performance measures with a focus on assessing and potentially correcting for selection bias in the literature, such as publication bias and *p*-hacking. First, a naïve random effects model was fit to the data and a moderate benefit of self-controlled practice, *g* = .44 (*k* = 52, *N* = 3134, 95% *CI* [.31, .56]), was found. Second, publication status was added to the model as a potential moderator, revealing a significant difference between published and unpublished findings, with only the former reporting a benefit of self-controlled practice. Third, to investigate and adjust for the impact of selectively reporting statistically significant results, a weight-function model was fit to the data with a one-tailed *p*-value cutpoint of .025. The weight-function model revealed substantial selection bias and estimated the true average effect of self-controlled practice as *g* = .107 (95% *CI* [.047, .18]). *P*-curve analyses were conducted on the statistically significant results published in the literature and the outcome suggested a lack of evidential value. Fourth, a suite of sensitivity analyses were conducted to evaluate the robustness of these results, all of which converged on trivially small effect estimates. Overall, our results suggest the benefit of self-controlled practice on motor learning is small and not currently distinguishable from zero.

@@

# NEED TO BE RUNNING THE DEVELOPMENT BRANCH VERSION OF PAPAJA
documentclass     : "apa7"
classoption       : "man, donotrepeattitle" # manuscript output
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage{ragged2e}
  - \usepackage{pdflscape}
  - \pagewiselinenumbers
  - \raggedbottom
  #- \leftheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  #- \rightheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  #- \hypersetup{draft} # comment out after but right now will prevent error message about some link spanning multiple pages
  #- \usepackage{setspace}
  #- \AtBeginEnvironment{longtable}{\singlespacing}
  #- \AtBeginEnvironment{tablenotes}{\doublespacing}
  #- \captionsetup[table]{font={stretch=1.5}}
appendix:
  - "appendix_a.Rmd"
  - "appendix_b.Rmd"
---

@@

Figure \@ref(fig:fig1) shows that the number of experiments comparing self-controlled groups to yoked groups has been increasing since the original experiments by Janelle and his colleagues [-@Janelle1995-rj; -@Janelle1997-ht]. Researchers have experimented with giving learners control over a variety of variables in the practice environment. A qualitative assessment of the literature suggests that self-control is generally beneficial regardless of choice-type [@Wulf2016-gf]. For example, self-control has been effective when participants have been provided choice over what can be considered instructionally-relevant variables, such as knowledge of results [@Patterson2010-tk], knowledge of performance [@Lim2015-qs], concurrent feedback [@Huet2009-wu], use of an assistive device [@Wulf2001-nb], observation of a skilled model [@Lemos2017-qx], practice schedule [@Wu2011-ha], practice volume [@Lessa2015-eq], and task difficulty [@Leiker2016-zd]. Additionally, self-controlled benefits have also been found for instructionally-irrelevant variables, such as: the colour of various objects in the practice environment [@Wulf2018-ie], other decorative choices [@Iwatsuki2019-tf], and the choice of what to do after the retention test is complete [@Lewthwaite2015-bd].

@@

In sum, the objectives of this meta-analysis were to estimate the true average effect of self-controlled learning and evaluate the evidential value of the self-controlled learning literature. Bias resulting from selective publication was addressed with weight function and *p*-curve models and effect size estimates were adjusted accordingly. A key theoretical question related to the underlying mechanisms of putative self-controlled learning advantages (motivational versus informational influences) was also addressed through moderator analyses, but, to anticipate, inferences will depend on the reliability of the evidence overall. Finally, sensitivity analyses were conducted in addition to pre-registered analyses in an effort to understand the extent that our conclusions depended on the modeling techniques and assumptions adopted.

@@

## Pre-registration
The procedures followed to conduct this meta-analysis were pre-registered and can be viewed at https://osf.io/qbg69 \hyperref[sec:sharing]{(see Data, materials, and code availability section)}. This meta-analysis was retrospective and earlier samples of the literature had been meta-analyzed prior to this pre-registration, albeit with different data collection procedures, scope, and excluding recent experiments. This study adheres to PRISMA reporting guidelines [@page2021].

## Literature search
The literature search and data extraction were conducted by three authors (BM, ZY, JH) and one research assistant (HS) independently. The goal of the search was to identify all articles that met the inclusion criteria for the meta-analysis. Specifically, randomized experiments were subject to five criteria for inclusion: 1) A self-control group in which participants were asked to make at least one choice during practice, 2) a yoked group that experienced the same practice conditions as the self-controlled group, 3) a delayed ~24-hour retention test or test with longer delay interval, 4) an objective measurement of motor performance, and 5) publication in a peer-reviewed journal or acceptance as part of a Master’s or PhD thesis. The literature search was completed on August 2, 2019.

The search commenced at PubMed and Google Scholar with the following query: *self-control** OR *self-regulat** OR *self-direct** OR *learner-control** OR *learner-regulat** OR *learner-direct** OR *subject-control** OR *subject-regulat** OR *subject-direct** OR *performer-control** OR *performer-regulat** OR *performer-direct** AND *motor learning**. The query retrieved 9014 hits on PubMed and 98,600 hits on Google Scholar. Each researcher excluded hits based on title alone or title and abstract when necessary, and quit searching the databases at self-selected intervals following extended periods of excluding 100% of search results. Following an initial run of searching databases, each researcher employed their own search strategies, including reviewing the reference sections of reviews and included articles, consulting the "OPTIMAL" theory website[^1], and searching the ProQuest Thesis database.

@@

Many of the self-controlled learning experiments analyzed in this study included multiple dependent measures. However, including multiple measures from the same experiment introduces bias and inflates Type 1 error [@Scammacca2014-do]. Although there are a variety of methods for dealing with multiple measures from the same studies in meta-analysis, we chose to create a priority list and always selected the highest priority dependent measure that was reported. If the highest priority measure was not described in adequate detail to calculate the effect size, the authors were contacted and the data were requested. If the authors could not provide the data for the highest priority dependent measure reported in their study, the experiment was left out of our analysis.

@@

## Data extraction
The four researchers separated into pairs and half of the included experiments were coded independently by one pair. The other half were coded independently by the other pair. The coding included varied moderators, publication year, and sample size. Also Hedges’ *g* was calculated from reported statistics and sample size using the `compute.es` package [@R-computees] in `R` [@R-base]. Effect sizes were calculated from means and standard deviations, test statistics like *t* and *F*, or from precisely reported *p*-values. When covariates were included in the analysis, the correlation coefficient for the covariate - dependent measure relationship was required to calculate accurate effect sizes. Since this information is often not reported, authors were contacted and the information was requested. One effect size was calculated for each of three time points for each experiment: Acquisition, retention, and transfer.

@@

The independent data extractions were compared and inconsistent results were highlighted. There was 89% absolute agreement between pairs of coders on 1344 data points. For those with disagreement, one of the researchers from the other coding pair reviewed the relevant experiment to confirm the value to be used in the analysis.[^3]

@@

In addition to extracting effect sizes, inferential statistics were scraped from published experiments that reported a statistically significant effect at retention. Two authors (BM and JH) independently completed a *p*-curve disclosure form consisting of a direct quote of the stated hypotheses for each experiment, the experimental design, and a direct quote of the results indicating a significant result (see Appendix A). There was 94% absolute agreement between the independent forms. Mismatches were resolved with consensus.

@@

## Risk of bias
All articles were assessed for risk of bias by the lead author using the Cochrane Risk of Bias 1.0 tool [@higgins2011]. Each article was coded as either high risk, unclear (some concerns), or low risk on 7 dimensions: sequence generation, allocation concealment, incomplete outcome data, selective outcome reporting, blinding of outcome assessment, blinding of participants and personnel, and other sources of bias.

@@

## Adjusting for selection effects
Selection bias in the motor learning literature is likely caused by filtering based on the statistical significance of results [@Lohse2016-cf]. To assess and adjust for selection effects, the `R` package `weightr` [@Coburn2017-hv]  was used to fit a Vevea-Hedges weight function model to the retention data [@Vevea1995-uv]. The weight-function model estimates the true average effect, heterogeneity, and the probability that a non-significant result survives censorship and is available for analysis. Selection effects are modelled by a step function that divides the effects into two bins at one-tailed *p* =.025, coinciding with a two-tailed *p*-value of .05. The probability of a non-significant effect surviving censorship to appear in the model is estimated relative to the probability of observing a study with a significant effect. The selection-adjusted model was compared to the naïve random effects model with a likelihood ratio test. Better fit from an adjusted model suggests selection bias in the literature.

@@

Bias correction methods vary in their performance depending on the total amount of heterogeneity, the true average effect size, the amount of publication bias, and the intensity of *p*-hacking in the data [@Carter2019-vv]. To determine which bias correction models perform well in the various plausible conditions for data in this meta-analysis, model performance checks were conducted using the \href{http://www.shinyapps.org/apps/metaExplorer/}{Meta-Showdown Explorer} shiny app developed by Carter and colleagues [-@Carter2019-vv]. Simulated conditions were as follows: Medium publications bias (significant results published at 100% frequency, non-significant published at 20% frequency, wrong direction effects published at 5% frequency), medium questionable research practice environment (QRP; see Carter et al. [-@Carter2019-vv] for detailed explanation of QRP environment), $\tau$ = 0, .2; *g* = 0, .2, .5; *k* = 60, good performance defined as a maximum of .1 upward or downward bias, and maximum mean absolute error of .1, also tested with maximum bias and error values of .15. With good performance defined by a maximum bias in either direction of .1 and maximum absolute error of .1, the weight function model and, to a lesser extent, *p*-curve models provided coverage across all plausible conditions except the highest heterogeneity condition ($\tau$ = .4). With good performance defined as a maximum bias and error of .15, the precision-effect with standard error (PEESE) method provided good performance in all conditions. Therefore, sensitivity analyses were conducted on effect size data via *p*-curve and PEESE methods. An additional sensitivity analysis of the estimated power among included studies was conducted with the *z*-curve [@bartovs2020]. *Z*-curve, like *p*-curve, analyzes only statistically significant results and estimates the power of the included studies (called expected replication rate, ERR). However, unlike *p*-curve, *z*-curve is robust to heterogeneity because it fits a finite mixture model of seven distributions, allowing the underlying true effects to vary. Further, *z*-curve also estimates the power of all studies that have been conducted (called expected discovery rate, EDR) which can be compared to the observed discovery rate in order to test for the presence of publication bias.

@@

## Risk of bias
The risk of bias assessment revealed lackluster reporting standards were pervasive among the included articles (see Figure \@ref(fig:fig2)). For example, comparing a self-control group to a yoked group usually involves first collecting a self-control participant, then their yoked counterpart. Despite this, most articles simply reported that the participants were randomly assigned to these conditions, with no indication of how this temporal constraint was addressed. A similar issue was observed with respect to addressing outliers and attrition. Over 75% of the included articles failed to mention outliers and how they were addressed (captured by the incomplete outcome data dimension). Most studies included in this study were not double-blind, largely due to the inherent difficulties in conducting a double-blind study of self-controlled motor learning. While the risk of bias associated with a lack of double blinding has been debated [see @howick2008], it is nonetheless notable that double-blinding was rare among the included studies.

(ref:fig2-caption) Proportion of studies with low risk, some concerns, and high risk of bias in each of the seven dimensions of the Cochrane RoB 1.0 tool.

@@

(ref:fig3-caption) Forest plot of Hedges' *g* (95% CI) for self-controlled versus yoked groups on retention tests. Size of squares is proportional to $1/\sigma^2$ (precision). Experiments are divided into published and unpublished subgroups and the black polygons represent 95% *CI* estimates from subgroup analyses. The black polygon at the bottom of the figure represents the 95% *CI* estimate for all included experiments.

```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center", out.height = "77%"}
knitr::include_graphics("../../figs/fig3.pdf")
```

@@

## Selection model
The weight-function model combines an effect size model and a selection model [@Hedges1996-yh]. The effect size model is equivalent to the naïve random effects model, specifying what the distribution of effect sizes would be in the absence of publication bias or other selection effects. The selection model accounts for the probability a given study survives selection based on its *p*-value and specifies how the effect size distribution is modified by selection. A weight-function model with a *p*-value cutpoint of (one-tailed) .025 was fit to the retention effect size estimates (see Figure \@ref(fig:fig4)). The results of a likelihood ratio test suggest the adjusted model was a significantly better fit to the data than the unadjusted model, $\chi^2$(*df* = 1) = 21.18, *p* < .0001.[^4] The adjusted effect size estimate was significantly different from zero, *g* = .107, *p* < .001, 95% *CI* [.05, .17]. According to the adjusted model, non-significant results were 6% as likely to survive selection as significant results. Note that the `weightr` function failed to estimate the random effects model and the results reported here are based on a fixed-effect estimate.

@@

(ref:fig4-caption) Funnel plot of self-controlled learning studies at retention. Standard error is plotted on the y-axis and Hedges' g is plotted on the x-axis. Dark gray contour regions represent two-tailed p-values between .10 and .05 (not quite significant). The light gray contour regions represent two-tailed p-values between .05 and .01. In the absence of bias (and other forms of heterogeneity), the most precise experiments would center on the naïve random effects estimate near the top of the plot and as experiments get progressively less precise they would move down the plot and spread out symmetrically. In the presence of bias, one would expect experiments to cluster in the light gray contour regions. The clustering of experiments in the positive light gray contour region in the above plot suggests substantial bias.

@@

## *P*-curve
The purpose of the *p*-curve analysis was to investigate the evidential value in the published reports (*N* = 26) of statistically significant self-controlled learning benefits. Visual inspection of Figure \@ref(fig:fig4) reveals a v-shaped distribution with the greatest frequency of *p*-values in the <.05 bin. The observed *p*-curve was significantly flatter than would be expected if the experiments had 33% power, *p* = .0035, indicating an absence of evidential value. Conversely, the half *p*-curve [@Simonsohn2015-mn] was significantly right skewed, suggesting the presence of evidential value. Sensitivity analysis, however, revealed that the half curve does not remain significantly right skewed following removal of the most extreme *p*-value from the sample. The estimated power of the included studies was 5%, 95% *CI* [5%, 17%].

## Interim discussion
The primary results described above suggest that selection effects have caused a seriously distorted record of self-controlled learning. Estimated benefits are less than one third of the naïve estimate, *g* = .107, 95% *CI* [.05, .17]. The *p*-curve analysis failed to detect robust evidence of a self-controlled learning effect. The performance of the weight-function model depends on the specific conditions present in the meta-analysis, although these conditions are unknowable [@Carter2019-vv]. It was necessary to conduct sensitivity analyses with additional bias correction methods to assess the reliability of the selection-adjusted weight-function model estimate. Based on performance checks conducted under a range of plausible conditions, it was determined that sensitivity analyses conducted with a PEESE meta-regression and *p*-curve effect size estimation would provide good performance coverage across most plausible conditions.

@@

## Precision-effect with standard error (PEESE) model
When publication bias is present in a body of evidence, sample size and effect size can be negatively correlated [@Stanley2014-xl]. The PEESE model fits a quadratic relationship between effect size and standard error to reflect the intuition that publication bias is stronger for low precision studies than high precision studies. The rationale is that low precision studies need to overestimate effects to achieve significance and get published, while high precision studies can publish without exaggerated effects; thus, creating greater publication bias among lower precision studies [@Carter2019-vv; @Stanley2014-xl]. A weighted-least-squares regression model was fit with effect size regressed on the square of the standard error, weighted by the inverse of the variance:
$$g_{i} = b_{0} + b_{1}se_{i}^2 + e_{i}$$
The PEESE method estimated a non-significant benefit of self-controlled learning after controlling for publication bias, *g* = .054, 95% *CI* [-.18, .29], *p* = .659.

## *P*-curve effect estimation
A *p*-curve model was fit to the overall retention effect size data, unlike the first primary *p*-curve which was fit to the reported significant results. The *p*-curve is a function of sample size and effect size, and because sample size is known, the effect size that provides the best fit to the observed *p*-curve can be estimated [@Simonsohn2014-xq]. A *p*-curve analysis conducted with th`R` package `dmetar` [@R-dmetar] was used to estimate the average effect size among the statistically significant effects in the meta-analysis. The model estimated an average effect of *g* = .035.[^5] The estimated power of included studies was 7%, 95% *CI* [5%, 22%]. Unfortunately, *p*-curve does not perform well in the presence of heterogeneity and these results should be interpreted cautiously.

@@

## *Z*-curve
A *z*-curve was fit to the overall retention data and estimated the power of statistically significant studies (ERR) as 14%, 95% CI [5%, 35%]. The power of all studies conducted (EDR) was estimated as 6%, 95% *CI* [5%, 14%]. The 95% confidence intervals for both the ERR and EDR failed to include the observed discovery rate of 48%, suggesting significant publication bias in the data.
